What We Eval in the Shadows

### 2021-04-06 analysis/environments.Rmd
Changes

  E_raw <- E_raw %>% mutate(envir_expression = if_else(envir_type == "NULL" & is.na(envir_expression), "NULL", envir_expression))

@@

```{r envir_default}
eval_calls_default <- E_envir %>% filter(is.na(envir))

@@

eval_calls_explicit <- E_envir %>% filter(!is.na(envir))
nb_explicit <- count(eval_calls_explicit, wt = duplicates)
nb_explicit_sites <- nb_eval_call_sites(eval_calls_explicit)

@@

```{r}
envir_var <- frequent_envir %>% filter(envir != "NA", str_detect(envir, "^(?:\\.[:alpha]|[:alpha:])[[:alpha:][:digit:]_\\.]*$"))

@@

```{r}
env_class_ranking <- E_envir %>% group_by(environment_class) %>% 
  summarize(n = n_distinct(src_ref)) %>% 
  arrange(desc(n))

@@

```{r}
frequent_envir %>% count(envir_type, sort = TRUE) %>% datatable()
```

@@

```{r}
E_envir %>% count(envir_type, sort = TRUE) %>% datatable()
```

@@

nb_global_env_sites <- specific_env %>% filter(specific_env == "global") %>% .$n
nb_callee_env_sites <- specific_env %>% filter(specific_env == "callee") %>% .$n
nb_empty_env_sites <- specific_env %>% filter(specific_env == "empty") %>% .$n
nb_base_env_sites <- specific_env %>% filter(specific_env == "base") %>% .$n

>>>>

  E_raw <- E_raw %>% mutate(envir_expression = if_else(envir_type == "NULL" & is.na(envir_expression), "NULL", envir_expression)) %>% 
    mutate(envir_expression = na_if(envir_expression, "NA"))

@@

```{r envir_default}
eval_calls_default <- E_envir %>% filter(is.na(envir)) # yes :(
# Or use E_envir %>% filter(ev_variant == "eval", !str_detect(ev_call, "\\benvir\\b|eval\\([^,]+,"))
# But that regex misses about 1000 sites

@@

eval_calls_explicit <- E_envir %>% filter(is.na(envir))
nb_explicit <- count(eval_calls_explicit, wt = duplicates)
nb_explicit_sites <- nb_eval_call_sites(eval_calls_explicit)

@@

```{r}
envir_var <- frequent_envir %>% filter(is.na(envir), str_detect(envir, "^(?:\\.[:alpha]|[:alpha:])[[:alpha:][:digit:]_\\.]*$"))

@@

```{r}
env_class_ranking <- E_envir %>% group_by(environment_class) %>% 
  summarize(n = n_distinct(src_ref), percent = round(100 * n / nb_call_sites, 2), cumpercent = cumsum(percent)) %>% 
  arrange(desc(n))

@@

```{r}
frequent_envir %>% count(envir_type, sort = TRUE) %>%
  mutate(percent = round(100 * n / nb_call_sites, 2), cumpercent = cumsum(percent)) %>% 
  datatable()
```

@@

```{r}
E_envir %>% count(envir_type, sort = TRUE)  %>%
  mutate(percent = round(100 * n / nb_call_sites, 2), cumpercent = cumsum(percent)) %>% 
 datatable()
```

@@

nb_global_env_sites <- specific_env %>% filter(specific_env == "global") %>% .$n
nb_callee_env_sites <- specific_env %>% filter(specific_env == "callee") %>% .$n
if(length(nb_callee_env_sites) == 0) nb_callee_env_sites <- 0
nb_empty_env_sites <- specific_env %>% filter(specific_env == "empty") %>% .$n
if(length(nb_empty_env_sites) == 0) nb_empty_env_sites <- 0
nb_base_env_sites <- specific_env %>% filter(specific_env == "base") %>% .$n
if(length(nb_base_env_sites) == 0) nb_base_env_sites <- 0

### 2021-04-06 paper/main.tex
More todo
Changes

\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}   %% For author/year citations
\usepackage{listings,hyperref,multirow,paralist,xspace,url,wrapfig}
\usepackage{tabularx}
\input{macro}
\newcommand{\mypara}[1]{\medskip\noindent\emph{#1}\xspace}
\newcommand{\NOTE}[1]{{\it Note: #1}\xspace}
\newcommand{\authorcomment}[3]{\xspace\textcolor{#1}{{\bf #2} #3}\xspace}
\newcommand{\todo}[1]{\authorcomment{red}{TODO}{#1}}

@@

\todo{Look at Improving dynamic code analysis by code abstraction, VPT'21}

@@

\begin{wrapfigure}{r}{5cm} \hspace*{-12mm}
  \centering
  \includegraphics[width=75mm]{pkgs-eval-callsites-hist.pdf} \caption{CRAN
  \eval call sites}%
  \label{fig:pkgs-eval-callsites-hist}
\end{wrapfigure}
%
This section focuses on CRAN packages and reports statistics about the usage of
\eval. There are \PkgEvalCallSites call sites to \eval in \PkgPackages
packages.  These represent \PkgPackagesRatio of all packages or \PkgCodeRatio
of the entire code size of CRAN. Over half of the packages have less than 3
\eval call sites and except \MaxEvalCallSitesPackage (\MaxEvalCallSitesCount
\eval calls), all packages contain less than \MaxEvalCallSitesRest \eval call
sites (\cf Fig~\ref{fig:pkgs-eval-callsites-hist}) These \eval calls appear in
\PkgFunsWithEval R functions (\CranFunsWithEvalRatio of all R functions in
CRAN).

For the dynamic analysis, we execute all the \CranRunnableScripts programs
extracted from \CranPackages packages. Any run that did not exercise one of the
\eval call sites in the packages was discarded. This left \Nbruns runs of some
of the \Corpus packages to generate the data presented here. There were
\Allcalls invocations of \eval and its variants occurring in \Triggeredpkgs
packages. Out of \PkgEvalCallSitesRnd \eval call sites, \PkgHitEvalCallSites
were hit at least once (\cf Fig~\ref{fig:traced-eval-callsites}). On average,
the analysis covers \PkgHitEvalCallSitesAvgRatio of package \eval call sites,
which similar to the \PkgCodeCoverage of average package code
coverage\footnote{Not all packages with static call sites of \eval were
observed triggering that call site, this can be chalked down to incomplete code
coverage and occasional tracer failure (\PkgFailedProgramsRatio of programs do
not terminate successfully or within 30 minute timeout).}.  Figure~\ref{freq}
summarizes the frequency of dynamic calls to \eval with the left column being
the number of calls and the right, the number of packages in that range. There
are \Fewcalls packages with low \eval frequency, fewer than 100 calls; and a
similar number of packages, \Manycalls to be precise, that use \eval more than
1,000 times.  One package, \Maxcallspack, calls it \Maxcalls times and thus
accounts for over half of the observed calls.

\begin{figure}[tb]
  \centering
  \includegraphics[width=.8\textwidth]{traced-eval-callsites.pdf}
  \centering
  \caption{\eval call sites coverage of the \PkgPackages packages.}%
  \label{fig:traced-eval-callsites}
\end{figure}

@@

To estimate how much executable code is injected through \eval, we measure the
the number of abstract syntax tree nodes as returned by the parser; for example,
\c{x+1} counts as 3.\footnote{We attempted to measure string lengths after
de-parsing expressions, but these measurements were dominated by de-parsed data
objects, which could range in the MBs.} The median argument size is
\Medianszeval (due to the many symbols) and the average is \Avgszeval nodes. The
largest \eval input observed is \Maxszeval\,-- a significant chunk of code.
Figure~\ref{fig:sizedistribution} shows the distribution of sizes for arguments
of fewer or equal to 25 nodes.
The x-axis is the size of arguments in number of
nodes, and the y-axis is the count of arguments with that size. The size drops
rapidly, with few observations larger than 15 nodes. The long tail is omitted
from the graph for legibility.
%
\begin{wrapfigure}{r}{5.8cm}
 \hspace*{-10mm}
 \includegraphics[width=0.5\textwidth, trim=4.8cm 0 6cm 0, clip]{size_loaded_distribution}
 \caption{Loaded code} \label{fig:sizedistribution}
\end{wrapfigure}
%

@@

\paragraph{Base}
We recorded \baseAllcalls \eval calls in the base libraries with fewer than 4\%
of those being to any variants. Unlike CRAN, Base calls \c{evalq} (\baseEvalqs).
Out of \baseStaticeval call sites, \baseTriggeredevalpct are exercised. Most
arguments, \baseCodepercent, are expressions. However, unlike CRAN,
\baseLanguagepercent of the arguments are language objects. The median argument
size is \baseMedianszeval, which is more than for CRAN. The maximum size is
\baseMaxszeval. Small instruction counts, less than 50, amount for
\baseSmalleventspct of calls.
\todo{More info about base and conclude it here.}

\paragraph{Kaggle}
Only \kaggleNbruns scripts trigger \eval. In total, \kaggleAllcalls \eval were
recorded; none are calls to variants. Out of \kaggleStaticeval call sites,
\kaggleTriggeredeval were exercised. Most call sites are invoked only once. Only
one site is called \kaggleMaxcalls times. Expressions amount for
\kaggleCodepercent of arguments. Unlike CRAN corpus, there are very few symbols
(\kaggleSymbolpercent). Most expressions result from calls to \c{parse}, thus
most \evals start with strings. The median argument size is \kaggleMedianszeval,
Which makes sense, as few arguments are only symbols. The largest argument is
\kaggleMaxszeval. The distribution of instructions per eval is similar to CRAN;
\kaggleSmalleventspct of \evals execute fewer than 50 instructions.
\todo{More info about kaggle and conclude it here.}

@@

\todo{Pierre: check the numbers - we want to know when it is an environment, when it is a list / data frame}
Most of the time, \ie in \packageEnvirEnvSitePercent of the sites, \eval is
passed an actual environment.

@@

% This table is old and incorrect
% Working on updating it
\begin{table}[htbp]{ \centering
\begin{tabular}{c|c|c|c}\hline
\multicolumn{2}{c|}{Core} & \multicolumn{2}{|c}{Packages}  \\\hline
Environment Class & Eval Calls \% & Environment Class & Eval Calls\\\hline
\CoreEnvClassA & \CoreProportionA & \PackagesEnvClassA &  \PackagesProportionA\\
\CoreEnvClassB & \CoreProportionB & \PackagesEnvClassB &  \PackagesProportionB\\
\CoreEnvClassC & \CoreProportionC & \PackagesEnvClassC &  \PackagesProportionC\\
\CoreEnvClassD & \CoreProportionD & \PackagesEnvClassD &  \PackagesProportionD\\
\CoreEnvClassE & \CoreProportionE & \PackagesEnvClassE &  \PackagesProportionE\\\hline
\end{tabular}} \label{tab:environments}
\caption{Environments in terms of \eval calls}
\end{table}

@@

\subsection{Provenance.}
% Explain tha some generate expression objects and other language objects?
There are various ways to obtain expressions that \eval will evaluate:
\begin{itemize}
    \item \c{substitute} synthesizes ASTs from expressions by replacing
    symbols with their bindings in the specified environment.
    \item \c{match.call} captures a call and its arguments as expressions.
    \item \c{quote}, \c{enquote} returns its argument unevaluated, as an R expression.
    \item \c{expression}  creates a vector of expression
    objects from text.
    \item \c{parse}, \c{str2expression} and \c{str2lang} turn strings into
    expressions.
    \item \c{comp} compiles R code into R bytecode.
\end{itemize}

\begin{table}[h]
  \begin{tabular}{|c|c|c|}
      \hline
      & \#sites & \%sites \\
       \hline
         \c{substitute}, \c{quote}, ... & \packageNbOtherLangsSites & \packageOtherLangsSitePercent \\
      \hline
      \c{parse}, \c{str2expression} & \packageNbParseExprsSites & \packageParseExprsSitePercent \\
      \hline
        \c{match.call} &  \packageNbMatchCallExprsSites & \packageMatchCallExprsSitePercent  \\   
      \hline
      \c{expression} & \packageNbNotParseExprsSites & \packageNotParseExprsSitePercent  \\
      \hline
      \c{str2lang} & \packageNbStrlangSites  & \packageStrlangSitePercent  \\
      \hline
      \c{comp} & \packageNbBytecodeSites & \packageBytecodeSitePercent \\
      \hline
  \end{tabular}
\caption{Provenances of \eval expressions. }
\end{table}

% \packageUnknownProvSitePercent : provenance not relavant (values, symbols etc)

\c{eval(parse(...))} can be used for dynamic code loading. This forms
the core of \c{source} and \c{sys.source} functions in R that are
commonly used for loading code in R files in interactive settings. We
investigated the number of cases in which the output of \c{parse} and
its variants is passed to \eval, directly or transitively by tainting
their output.  We observed that very few of the eval call (\packageNbParseFromFileSites in total)
consume the result of calling \c{parse} on a file. Most of the eval
calls consume the result of calling \c{parse} on a string. We also
identified one function in core R, \c{invokeRestartInteractively} that
prompts the user for input, parses it, and passes it to \eval.

\subsection{Side effects in \eval}

\todo{Filip: discuss the side-effects}

@@

\todo{Illustrate things that we have discussed in Section 5 - examples}
\todo{Things to discuss: model.frame, match.call, substitute, parse, plots}

% We present here the main uses of \eval we have witnessed in R. Sometimes, they are design patterns, sometimes, \eval could be replaced by more specific functions. For each use of \eval, we discuss whether it is necessary to use.

% \subsection{Variable lookup}

% \eval is used to look-up the value of a variable in a given environment, different from the current scope so the \c{envir} argument of \eval is explicitly specified.

% \subsection{Metaprogramming}

% \mypara{\c{substitute}}

% \mypara{\c{match.call}}

% \todo{Talk about model.frame}

% \subsection{Logging}

% \subsection{Code transform}

% % Derivation of symbolic expressions

% \subsection{Plotting}

% \section{\eval expressivity and macros}

@@

>>>>

\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}   %% For author/year citations
\usepackage{listings,hyperref,multirow,paralist,xspace,url,wrapfig}
\usepackage{tabularx}
\input{macro}
\newcommand{\mypara}[1]{\medskip\noindent\emph{#1}\xspace}
\newcommand{\NOTE}[1]{{\it Note: #1}\xspace}

@@

@@

\begin{wrapfigure}{r}{5cm} \hspace*{-12mm}
  \includegraphics[width=75mm]{pkgs-eval-callsites-hist.pdf} \caption{CRAN
  \eval call sites}%
  \label{fig:pkgs-eval-callsites-hist}
\end{wrapfigure}
%
This section focuses on CRAN packages and reports statistics about the usage of
\eval. There are \PkgEvalCallSites calls to \eval in \PkgPackages packages.
These represent \PkgPackagesRatio of all packages or \PkgCodeRatio of the
entire code size of CRAN. Over half of the packages have less than 3 \eval
call sites and except \MaxEvalCallSitesPackage (\MaxEvalCallSitesCount \eval
calls), all packages contain less than \MaxEvalCallSitesRest \eval call sites
(\cf Fig~\ref{fig:pkgs-eval-callsites-hist}) These \eval calls appear in
\PkgFunsWithEval R functions (\CranFunsWithEvalRatio of all R functions in
CRAN).

For the dynamic analysis, we execute all the \CranRunnableScripts programs
extracted from \CranPackages packages. Any run that did not exercise one of the
\eval call sites in the packages was discarded. This left \Nbruns runs of some
of the \Corpus packages to generate the data presented here. There were
\Allcalls invocations of \eval and its variants occurring in \Triggeredpkgs
packages. Out of \PkgEvalCallSitesRnd \eval call sites, \PkgHitEvalCallSites
were hit at least once\footnote{Not all packages with static call sites of
\eval were observed triggering that call site, this can be chalked down to
incomplete code coverage and occasional tracer failure (\PkgFailedProgramsRatio
of programs do not terminate successfully or within 30 minute timeout).}. On
average, the analysis covers \PkgHitEvalCallSitesAvgRatio of package \eval call
sites, which similar to the \PkgCodeCoverage of average package code coverage.
Figure~\ref{freq} summarizes the frequency of dynamic calls to \eval with the
left column being the number of calls and the right, the number of packages in
that range. There are \Fewcalls packages with low \eval frequency, fewer than
100 calls; and a similar number of packages, \Manycalls to be precise, that use
\eval more than 1,000 times.  One package, \Maxcallspack, calls it \Maxcalls
times and thus accounts for over half of the observed calls.

@@

To estimate how much executable code is injected through \eval, we measure the
the number of abstract syntax tree nodes as returned by the parser; for example,
\c{x+1} counts as 3.\footnote{We attempted to measure string lengths after
de-parsing expressions, but these measurements were dominated by de-parsed data
objects, which could range in the MBs.} The median argument size is
\Medianszeval (due to the many symbols) and the average is \Avgszeval nodes. The
largest \eval input observed is \Maxszeval\,-- a significant chunk of code.
Figure~\ref{fig:sizedistribution} shows the distribution of sizes for arguments
of fewer or equal to 25 nodes.
%
\begin{wrapfigure}{r}{5.8cm}
 \hspace*{-10mm}
 \includegraphics[width=0.5\textwidth, trim=4.8cm 0 6cm 0, clip]{size_loaded_distribution}
 \caption{Loaded code} \label{fig:sizedistribution}
\end{wrapfigure}
%
The x-axis is the size of arguments in number of
nodes, and the y-axis is the count of arguments with that size. The size drops
rapidly, with few observations larger than 15 nodes. The long tail is omitted
from the graph for legibility.

@@

\paragraph{Base}
We recorded \baseAllcalls \eval calls in the base libraries with fewer than 4\%
of those being to any variants. Unlike CRAN, Base calls \c{evalq} (\baseEvalqs).
Out of \baseStaticeval call sites, \baseTriggeredevalpct are exercised. Most
arguments, \baseCodepercent, are expressions. However, unlike CRAN,
\baseLanguagepercent of the arguments are language objects. The median argument
size is \baseMedianszeval, which is more than for CRAN. The maximum size is
\baseMaxszeval. Small instruction counts, less than 50, amount for
\baseSmalleventspct of calls.

\paragraph{Kaggle}
Only \kaggleNbruns scripts trigger \eval. In total, \kaggleAllcalls \eval were
recorded; none are calls to variants. Out of \kaggleStaticeval call sites,
\kaggleTriggeredeval were exercised. Most call sites are invoked only once. Only
one site is called \kaggleMaxcalls times. Expressions amount for
\kaggleCodepercent of arguments. Unlike CRAN corpus, there are very few symbols
(\kaggleSymbolpercent). Most expressions result from calls to \c{parse}, thus
most \evals start with strings. The median argument size is \kaggleMedianszeval,
Which makes sense, as few arguments are only symbols. The largest argument is
\kaggleMaxszeval. The distribution of instructions per eval is similar to CRAN;
\kaggleSmalleventspct of \evals execute fewer than 50 instructions.

@@

Most of the time, \ie in \packageEnvirEnvSitePercent of the sites, \eval is
passed an actual environment.

In \packageNbEnvirIncorrectSites sites (\packageEnvirIncorrectSitePercent),
\eval is passed an incorrect argument, for instance \c{NA}, a boolean, which
makes \eval fail.

@@

% This table is old and incorrect
% Working on updating it
\begin{table}[htbp]{ \centering
\begin{tabular}{|c|c|}\hline
 Environment Class & Eval Calls\\\hline
 \PackagesEnvClassA &  \PackagesProportionA\\
 \PackagesEnvClassB &  \PackagesProportionB\\
 \PackagesEnvClassC &  \PackagesProportionC\\
 \PackagesEnvClassD &  \PackagesProportionD\\
\PackagesEnvClassE &  \PackagesProportionE\\\hline
\end{tabular}} \label{tab:environments}
\caption{Environments in terms of \eval calls}
\end{table}

@@

\subsection{Provenance.} 
% Explain that some generate expression objects and other language objects?
There are various ways to obtain expressions that \eval will evaluate:
\begin{itemize}
    \item \c{substitute} synthesizes ASTs from expressions by replacing
    symbols with their bindings in the specified environment.
    \item \c{match.call} captures a call and its arguments as expressions.
    \item \c{quote}, \c{enquote} returns its argument unevaluated, as an R expression. % Aldd \c{call} ?
    \item \c{expression}  creates a vector of expression
    objects from text.
    \item \c{parse}, \c{str2expression} and \c{str2lang} turn strings into
    expressions.
    \item \c{comp} compiles R code into R bytecode.
\end{itemize}
We show the distribution of the different provenances in Table~\ref{tab:provenance}. Note that one site could be passed arguments from different provenances depending on the invokation. For some arguments, we could not determine their provenance or it was a value or a symbol, for which the provenance is less significant. % Not clear... Or maybe add symbols in that table?
% One can create a symbol with as.name, as.symbol etc. That would be the provenance (but a symbol can also come from quote for instance...)

\begin{table}[h]
    \begin{tabular}{|c|c|c|}
        \hline
        & \#sites & \%sites \\
        \hline
        \c{substitute}, \c{quote}, ... & \packageNbOtherLangsSites & \packageOtherLangsSitePercent \\
        \hline
        \c{parse}, \c{str2expression} & \packageNbParseExprsSites & \packageParseExprsSitePercent \\
        \hline
        \c{match.call} &  \packageNbMatchCallExprsSites & \packageMatchCallExprsSitePercent  \\   
        \hline
        \c{expression} & \packageNbNotParseExprsSites & \packageNotParseExprsSitePercent  \\
        \hline
        \c{str2lang} & \packageNbStrlangSites  & \packageStrlangSitePercent  \\
        \hline
        \c{comp} & \packageNbBytecodeSites & \packageBytecodeSitePercent \\
        \hline \hline 
        Other & \packageNbUnknownProvSites & \packageUnknownProvSitePercent \\
        \hline
    \end{tabular}
    \caption{Provenances of \eval expressions. } \label{tab:provenance}
\end{table}

We taint the result of a call to \c{parse}, \c{str2lang} and \c{str2expression}, by attaching an R attribute to it. R attributes are lost under copy or modification of their attaching object, so we underestimate the number of sites in that case. However, we observed that most of the time, the result of \c{parse} is not touched and directly passed to \eval.

We taint the result of \c{match.call} by remembering the addresses of all the elements of the call list. As R GC is non-moving, it makes it possible to track them individually, as we observed that calls are modified and individual elements are sometimes evaluated. This is also an underestimation of the number of sites though, as we would lose an individual element of the call list that would be modified.

% explain also how we get the numbers of the other categories?

\c{eval(parse(...))} (and \c{str2expression}, and \c{str2lang}) can be used for dynamic code loading. This forms
the core of \c{source} and \c{sys.source} functions in R that are
commonly used for loading code in R files in interactive settings. We
investigated the number of cases in which the output of \c{parse} and
its variants is passed to \eval, directly or transitively by tainting
their output.  We observed that very few of the eval call (\packageNbParseFromFileSites in total)
consume the result of calling \c{parse} on a file. Most of the eval
calls consume the result of calling \c{parse} on a string. We also
identified one function in core R, \c{invokeRestartInteractively} that
prompts the user for input, parses it, and passes it to \eval.

In Kaggle, most of the sites, \ie \kaggleParseExprsSitePercent, are passed  an expressions that result from a call to \c{parse}.

\mypara{New Bindings.} Find number of times, eval introduces new bindings. There
are many ways -- library, load, attach, source, and explicitly introduce
bindings using super-assign, assign and define.

\mypara{C Calls.} Find number of times, eval makes call to native code using -
.Call, .External, etc. Find out if there are native functions that are called
only from within eval.

\mypara{Non-local returns} Eval can do non-local returns effective bypassing
evaluation of the rest of the function. For instance, it can inject a \c{break} or \c{next} instructions in a loop. It is very rare however, as it happens only twice in our dataset. % What kind of non-local returns?
% eval(quote(return(89))) does not work

\mypara{Purity.} Conclude with number of evals that are ``pure'', i.e. evals
which could be ignored by a static analyzer without any problem.

@@

We present here the main uses of \eval we have witnessed in R. Sometimes, they are design patterns, sometimes, \eval could be replaced by more specific functions. For each use of \eval, we discuss whether it is necessary to use.

\subsection{Variable lookup}

\eval is used to look-up the value of a variable in a given environment, different from the current scope so the \c{envir} argument of \eval is explicitly specified.

% TODO: numbers on the prevalence of the pattern

\eval is not necessary in that case and can be replaced by \c{get}, as follows:

\begin{lstlisting}
 eval(quote(x), envir = env)
\end{lstlisting}
$\Downarrow$
\begin{lstlisting}
get("x", envir = env)
\end{lstlisting}

% This is actually linked to NSE, as the symbol given to \eval usually comes
% from match.call or susbtitute

\subsection{Non-Standard Evaluation} % Or metaprogramming

\subsection{\c{substitute}}

\subsection{\c{match.call}}

\subsection{Logging}

\subsection{Code transform}

% Derivation of symbolic expressions

\subsection{Plotting}

\section{\eval expressivity and macros}

@@

\section{The dangers of \eval} \label{sec:dangerseval}

\subsection{Eval Usage in R}
